---
title: "GPR - example 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gpr_ex1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
        echo = TRUE, results = 'hold', warning=F, cache=F, 
  #dev = 'pdf', 
  message=F, 
  fig.width=5, fig.height=5,
  tidy.opts=list(width.cutoff=75), tidy=FALSE
)
options(scipen = 1, digits = 4)
```

```{r setup}
library(GPFDA)
require(MASS)
```

## Simulating data from a GP with unidimensional input

We simulate $30$ independent realisations from a zero-mean GP with a covariance 
function given by the sum of a liner kernel and a squared exponential kernel. 
Each observed curve has a sample size of $15$ time points on $[0,1]$.

```{r}
set.seed(123)
nrep <- 30
n <- 15
input <- seq(0, 1, length.out=n)
hp <- list('linear.a'=log(40), 'linear.i'=log(10),
           'pow.ex.v'=log(5), 'pow.ex.w'=log(15),
           'vv'=log(0.3))
Sigma <- cov.linear(hyper=hp, input=input) + 
  cov.pow.ex(hyper=hp, input=input, gamma=2) + 
  diag(exp(hp$vv), n, n)
Y <- t(mvrnorm(n=nrep, mu=rep(0,n), Sigma=Sigma))
```

## Estimation

Estimation of the GPR model can be carried out without using gradient:
```{r}
set.seed(111)
fitNoGrad <- gpr(input=input, response=Y, Cov=c('linear','pow.ex'), gamma=2, 
               trace=4, nInitCandidates = 1, useGradient = F)
```

If one wants to use gradient:
```{r}
set.seed(111)
fit <- gpr(input=input, response=Y, Cov=c('linear','pow.ex'), gamma=2, 
         trace=4, nInitCandidates = 1, useGradient = T)
```

Note the smaller number of iterations needed when the gradient analytical 
expressions are used in the optimisation.

We can see that the hyperparameter estimates are very accurate despite the fairly 
small sample size:
```{r}
sapply(fit$hyper, exp)
```

The fitted model for the $10$th realisation can be seen:
```{r}
plot(fit, realisation=10)
```


## Prediction

Predictions on a fine grid for the $10$th realisation can be obtained as follows:
```{r}
input.new <- seq(0, 1, length.out = 1000)
pred1 <- gprPredict(train=fit, input.new=input.new, noiseFreePred=T)
plot(pred1, realisation=10)
```

If one wants to include noise variance in the predictions for the new time points:
```{r}
pred2 <- gprPredict(train=fit, input.new=input.new, noiseFreePred=F)
plot(pred2, realisation=10)
```